Consumption and resistance across states and HRRs

### 2017-03-01 states/resistance/state-report.Rmd
moved to filter

`r data_frame(organsisms=organisms) %>% kable`

@@

```{r load_data}

coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state=level %in% c('state', 'hrr', 'zipcode'),
         hrr=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state, hrr) %>%
  gather('place', 'coherent', state, hrr)

hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

@@

```{r run_models}
safe_spearman = function(df, expl, indep) {
  if (dim(df)[1] < 3L) {
    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
  } else {
    cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy
  }
}

@@

models_result = function(x, this_place, consumption) {
  result = x %>%
    group_by_('bug', 'resistance_drug', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE)) %>%
    left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
    left_join(consumption, by=c(this_place, 'consumption_drug')) %>%
    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
    group_by(bug, resistance_drug, consumption_drug) %>%
    do(models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
    mutate(record_number=1:n(),
           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
           plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
    select(bug, consumption_drug, resistance_drug, coherent,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig,
           plot_fn, plot_link)

  # add adjusted p-values for just the coherent stuff
  # swo> there's probably a more elegant way to do this
  coherent.adj.p.value = result %>%
    filter(coherent) %$%
    p.adjust(p.value, method='BH')
  coherent.sig = pvalue_stars(coherent.adj.p.value)
  result$coherent.adj.p.value = NA
  result$coherent.sig = NA
  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
  result$coherent.sig[result$coherent] = coherent.sig

  result
}

state_res = models_result(state_dat, 'state', state_consumption)
hrr_res = models_result(hrr_dat, 'hrr', hrr_consumption)

@@

  p = plot_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug) %>%
    ggplot(aes_string(row$explanatory, 'percent_nonsusceptible')) +
    geom_point(shape=1) +
    xlab(row$explanatory) +
    ylab('percent non-susceptible') +
    theme_minimal()

@@

### All resistance-averaged results

@@

#### All state-level averaged-resistance correlations

```{r state_average_results_table}
result %>%
  filter(place=='state') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

There correlations are mostly positive, but only a small number are significant (red line marks $p = 0.05$). The significant negative correlation is for MRSA and TMP-SMX.

@@

#### All HRR-level averaged-resistance

```{r hrr_average_results_table}
result %>%
  filter(place=='hrr') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

Like with the state-level data, only a small number are significant (red line marks $p = 0.05$), but the correlations seem more split positive and negative.

@@

The results from the HRR level differ: both identified a relationship between *E. coli* and cipro and MSSA and azithromycin. However, the state-level data  identified a negative relationship between MRSA and Bactrim while the HRR data identified a negative relationship between *K. pneumoniae* and Bactrim. The state-level data also identified *S. pneumoniae* and azithromycin.

### Coherent resistance-averaged

Second, I show the results from the coherent tests only. The $p$-values here are different because I adjusted those values using only this subset of tests.

The plots (in the links) here are the same as before. They're actually showing all the data. I should probably have plots that show just the averaged points.

#### Coherent state-level averaged-resistance correlations

```{r coherent_state_average_results_table}
result %>%
  filter(place=='state', coherent) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         explanatory, correlation,
         coherent.adj.p.value, coherent.sig,
         plot_link) %>%
  kable
```

There correlations are mostly positive, and one or two more are significant (red line marks $p = 0.05$):

```{r coherent_state_average_histogram}
result %>%
  filter(place=='state', coherent) %>%
  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('correlation coefficient') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

#### All HRR-level averaged-resistance

```{r coherent_hrr_average_results_table}
result %>%
  filter(place=='hrr', coherent) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         explanatory, correlation,
         coherent.adj.p.value, coherent.sig,
         plot_link) %>%
  kable
```

```{r coherent_hrr_average_histogram}
result %>%
  filter(place=='hrr', coherent) %>%
  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('correlation coefficient') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

### All bug/drug, all-antibiogram results

@@

```{r}
all_models_result = function(x, this_place, consumption) {
  result = x %>%
    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
    group_by(bug, resistance_drug, consumption_drug) %>%
    do(models(., 'percent_nonsusceptible')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    left_join(filter(coherence, place==this_place), by=c('bug', 'resistance_drug')) %>%
    select(bug, consumption_drug, resistance_drug, coherent,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig)

  # add adjusted p-values for just the coherent stuff
  # swo> there's probably a more elegant way to do this
  coherent.adj.p.value = result %>%
    filter(coherent) %$%
    p.adjust(p.value, method='BH')
  coherent.sig = pvalue_stars(coherent.adj.p.value)
  result$coherent.adj.p.value = NA
  result$coherent.sig = NA
  result$coherent.adj.p.value[result$coherent] = coherent.adj.p.value
  result$coherent.sig[result$coherent] = coherent.sig

  result
}

@@

#### All bug/drug, state-wide consumption, all-antibiogram

@@

#### All bug/drug, HRR-wide consumption, all-antibiogram

@@

#### Coherent, state-wide consumption, all-antibiogram

```{r}
all_result %>%
  filter(place=='state', coherent) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         explanatory, correlation,
         coherent.adj.p.value, coherent.sig) %>%
  kable
```

```{r}
all_result %>%
  filter(place=='state', coherent, explanatory=='cpkp') %>%
  ggplot(aes(x=correlation, y=coherent.adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  scale_y_reverse() +
  theme_minimal()
```

#### Coherent, HRR-wide consumption, all-antibiogram

```{r}
all_result %>%
  filter(place=='hrr', coherent) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         explanatory, correlation,
         coherent.adj.p.value, coherent.sig) %>%
  kable
```

>>>>

`r organisms`

@@

I used the 2011 Medicare data.

```{r load_data}
hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

@@

```{r run_models}
safe_spearman = function(df, expl, indep) {
  try(return(cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy), silent=TRUE)
  data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
}

@@

models_result = function(x, this_place) {
  x %>%
    group_by_('bug', 'drug_group', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
              cpkp=head(cpkp, 1),
              did=head(did, 1)) %>%
    group_by(bug, drug_group) %>%
    do(models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    mutate(record_number=1:n(),
           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
           plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
    select(bug, drug_group,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig,
           plot_fn, plot_link)
}

state_res = models_result(state_dat, 'state')
hrr_res = models_result(hrr_dat, 'hrr')

@@

  p = plot_dat %>%
    filter(bug==row$bug, drug_group==row$drug_group) %>%
    ggplot(aes_string(row$explanatory, 'percent_nonsusceptible')) +
    geom_point(shape=1) +
    xlab(row$explanatory) +
    ylab('percent non-susceptible') +
    theme_minimal()

@@

### Resistance-averaged results

@@

#### State-level resistance-averaged correlations

```{r state_average_results_table}
result %>%
  filter(place=='state') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

(red line marks $p = 0.05$).

@@

#### HRR-level resistance-averaged

```{r hrr_average_results_table}
result %>%
  filter(place=='hrr') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

(red line marks $p = 0.05$)

@@

**swo adjust** The results from the HRR level differ: both identified a relationship between *E. coli* and cipro and MSSA and azithromycin. However, the state-level data  identified a negative relationship between MRSA and Bactrim while the HRR data identified a negative relationship between *K. pneumoniae* and Bactrim. The state-level data also identified *S. pneumoniae* and azithromycin.

### All-antibiogram results

@@

```{r}
all_models_result = function(x, this_place, consumption) {
  x %>%
    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
    group_by(bug, drug_group) %>%
    do(models(., 'percent_nonsusceptible')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    select(bug, drug_group,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig)
}

@@

#### State-wide consumption, all-antibiogram

@@

#### HRR-wide consumption, all-antibiogram

@@
